Human tissue specific markers

### 2016-12-21 manuscript/manuscript.Rnw
switch to acro from acronym
work on manuscript

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{txfonts}
\usepackage{fancyref}
\usepackage{acronym}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,style=numeric-comp]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{xspace}
% this package will stretch the font slightly if necessary.
\usepackage{microtype}
\usepackage{zref-xr,zref-user}
\newenvironment{narrow}[2]{%
  \begin{list}{}{%
      \setlength{\topsep}{0pt}%
      \setlength{\leftmargin}{#1}%
      \setlength{\rightmargin}{#2}%
      \setlength{\listparindent}{\parindent}%
      \setlength{\itemindent}{\parindent}%
      \setlength{\parsep}{\parskip}}%
  \item[]}{\end{list}}
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
\pagestyle{fancy}

@@

<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs","svm_test_results","svm_test_results_reduced")))>>=
for (a in paste0("../tissue_specific_expression/",
                 c("tissue_specific_markers",
                   "interesting_gene_reads_wide",
                   "combined_star_logs",
                   "svm_test_results",
                   "svm_test_results_reduced"
                   ))) {
    load(a)
}
tissue.specific.genes <- data.table(tissue.specific.genes)
setkey(tissue.specific.genes,"gene")
@

@@

\acp{EV} are small (40-1000nm) compartments that can be found in
blood, saliva, and urine. \acp{EV} originate from the plasma membrane
or multivesicular endosome of many cell types and contain proteins,
lipids, mRNA, miRNA, and many other non-coding RNAs. Because \acp{EV}
contain mRNA from the cell of origin and can be obtained without
invasive procedures, they are a promising remote sensor of the
transcriptome of tissues that would otherwise be inaccessible.

@@

@@

@@

\subsection{Uterus Specific}
<<uterus.specific.markers.table,results="asis",echo=FALSE>>=
temp <- tissue.specific.genes[tissue=="uterus",
                              list(gene,tissue,specificity,gi.max)]
setorder(temp,-gi.max)
setnames(temp,c("Gene","Tissue","Tissue Specificity","Mean Expression"))
print(xtable(temp,
             caption="Uterus specific genes",label="tab:uterus_specific_genes"),
      include.rownames=FALSE,
      )
rm(temp)
@

@@

\subsection{Placenta Specific Genes}

@@

\subsection{Identification of Tissue source of Exosomes}

@@

\begin{figure}
\label{fig:plac_ut_red_roc}
\end{figure}

>>>>

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{txfonts}
\usepackage{fancyref}
\usepackage{acro}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,style=numeric-comp]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{xspace}
% this package will stretch the font slightly if necessary.
\usepackage{microtype}
\usepackage{zref-xr,zref-user}
\usepackage[capitalize]{cleveref}
\newenvironment{narrow}[2]{%
  \begin{list}{}{%
      \setlength{\topsep}{0pt}%
      \setlength{\leftmargin}{#1}%
      \setlength{\rightmargin}{#2}%
      \setlength{\listparindent}{\parindent}%
      \setlength{\itemindent}{\parindent}%
      \setlength{\parsep}{\parskip}}%
  \item[]}{\end{list}}
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
\pagestyle{fancy}

\newcommand{\EV}{\EV}

@@

<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=tools::md5sum(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs","svm_test_results","svm_test_results_reduced","housekeeping_genes_superset")))>>=
for (a in paste0("../tissue_specific_expression/",
                 c("tissue_specific_markers",
                   "interesting_gene_reads_wide",
                   "combined_star_logs",
                   "svm_test_results",
                   "svm_test_results_reduced",
                   "housekeeping_genes_superset"
                   ))) {
    load(a)
}
tissue.specific.genes <- data.table(tissue.specific.genes)
setkey(tissue.specific.genes,"gene")
@

@@

\subsection{Housekeeping genes}

\subsection{Cell free RNA and Exosomes in Diagnostics}

\EV are small (40-1000nm) compartments that can be found in
blood, saliva, and urine. \EV originate from the plasma membrane
or multivesicular endosome of many cell types and contain proteins,
lipids, mRNA, miRNA, and many other non-coding RNAs. Because \EV
contain mRNA from the cell of origin and can be obtained without
invasive procedures, they are a promising remote sensor of the
transcriptome of tissues that would otherwise be inaccessible.

@@

\subsection{Tissue RNA-seq}

\subsection{Tissue Specificity}

@@

\subsection{Read Biasing}

The population of RNAs found in \EV are distinct from the cytosolic
pool, and are likely selectively sorted into \EV~%
\cite{Villarroya-Beltri.ea2014:Sortingitoutregulation,Hoen.ea2012:DeepsequencingofRNA,Janas.ea2015:MechanismsofRNAloading,Balkom.ea2015:Quantitativeandqualitativeanalysis}.
Specifically, the motives
GGAG~\cite{Villarroya-Beltri.ea2013:SumoylatedhnRNPA2B1controlssorting},
ACCAGCCU, CAGUGAGC, and
UAAUCCCA~\cite{Batagov.ea2013:Exosomessecretedbyhuman} are enriched in
\EV. Owing to the limited number of tissue-specific \ac{EV} RNA-seq
results, we have biased the reads from existing tissue-specific
RNA-seq results to preferentially include RNAs which contain motifs
found in exosomes and also match the lower number of unique RNAs which
are found in individual exosomes.

\subsection{Machine Learning}

@@

\subsection{Tissue Specific Genes and Housekeeping genes}

\begin{figure}
<<tissue.specificity,shape.figure,fig.width=8,fig.height=8,echo=FALSE>>=
invisible(tissue.specific.genes[,type:="not specific"])
invisible(tissue.specific.genes[,housekeeping:="not housekeeping"])
invisible(tissue.specific.genes[possibly.specific==TRUE,type:="tissue specific"])
invisible(tissue.specific.genes[tracking_id %in% housekeeping.genes.superset[,tracking_id] &
                                 possibly.specific==FALSE,housekeeping:="housekeeping"])
invisible(tissue.specific.genes[tracking_id %in% housekeeping.genes.superset[,tracking_id] &
                          possibly.specific==TRUE,housekeeping:="housekeeping and specific"])

invisible(tissue.specific.genes[,type:=
                          factor(type,
                                 levels=c("not specific","tissue specific",
                                     "housekeeping",
                                     "housekeeping and specific"
                                          ))])
print(ggplot(tissue.specific.genes[gi.max>0],
             aes(y=specificity,x=gi.max,colour=type))+
      ##geom_point()+
      geom_hex(bins=50)+
      ## stat_summary_2d(fun=function(x){length(x)},bins=100)+
      geom_point(aes(colour=housekeeping),data=tissue.specific.genes[housekeeping!="not housekeeping",])+
      scale_x_log10()+
      xlab("Max Expression (FPKM)")+
      ylab("Tissue Specificity")+
      theme(legend.position="top")
    )
@ 
\caption{Shape of tissue specific genes}
\label{fig:tissue_specific_shape}
\end{figure}

\Cref{fig:tissue_specific_shape} shows the shape of the tissue
specificity measure of all genes along with the maximum expression in
a particular tissue. Almost all housekeeping genes identified by
\citet{Eisenberg.ea2013:Humanhousekeepinggenesrevisited} have
$\tau_g \le 0.93$ with the exception of
%
\Sexpr{tissue.specific.genes[housekeeping=="housekeeping and specific",gene]}
%

\subsection{Uterus and Placenta Specific Genes}

<<uterus.specific.markers.table,results="asis",echo=FALSE>>=
temp <- tissue.specific.genes[tissue=="uterus",
                              list(gene,tissue,specificity,gi.max)]
setorder(temp,-gi.max)
setnames(temp,c("Gene","Tissue","Tissue Specificity","Mean Expression"))
print(xtable(temp,
             caption="Uterus specific genes",
             label="tab:uterus_specific_genes"),
      include.rownames=FALSE,
      )
rm(temp)
@

@@

@@

\subsection{Biasing reads towards exosomes}

\subsection{Machine learning to identify tissue source}

@@

<<calc.reduced,echo=FALSE,results="asis">>=
reduced.reads <- function(fpkms,num.reads) {
    non.zero <- fpkms > 0
    ret.val <- rep.int(0,length(fpkms))
    if (!any(non.zero)) {
        return(fpkms)
    }
    inc.prob <- inclusionprobabilities(fpkms[non.zero],1)
    ret.val[non.zero] <-
        apply(sapply(1:(num.reads/1),function(y){UPrandomsystematic(inc.prob,0)}),1,sum)
    names(ret.val) <- names(fpkms)
    ret.val
}
set.seed(1021)
reads <- c("A"=1000,
           "B"=2000,
           "C"=20,
           "D"=5000)
reads.to.table <- function(read,reads){
    data.table("Total Reads"=read,
               "Gene A"=reads["A"],
               "Gene B"=reads["B"],
               "Gene C"=reads["C"],
               "Gene D"=reads["D"])
}
rr.table <- list(reads.to.table("Orig. FPKM",reads))
i <- 0
for (read in c(500,500,100,100,5,5,5)){
    i <- i+1
    reads <- reduced.reads(reads,read)
    rr.table[[paste0("read.",read,'.',i)]] <- 
        reads.to.table(read,reads)
}
print(xtable(rbindlist(rr.table),
             digits=0
             ),
      include.rownames=FALSE)
@
